### 2021-04-24 submission/manuscript.Rmd
Edits and additions to results

---
title: ''
csl: mbio.csl #Get themes at https://github.com/citation-style-language/styles
fontsize: 11pt
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
  word_document: default
  html_document:
    df_print: paged
geometry: margin=1.0in
bibliography: references.bib
---

@@

**5-day laxative treatment leads to prolonged *C. difficile* colonization in mice.** We compared mice treated with the osmotic laxative PEG 3350 to our standard 10 mg/kg clindamycin treatment, which temporarily renders the mice susceptible to *C. difficile*, with mice typically clearing *C. difficile* within 10 days post-infection [@Tomkovich2019; @Tomkovich2020]. All PEG-treated mice were administered a 15% PEG solution in the drinking water for 5-days, one group was also treated with clindamycin, and one group was allowed to recover for 10 days prior to challenge (Fig. 1A). PEG treatment resulted in weight loss in all 3 groups of PEG-treated mice, with the greatest change in weight observed on the fifth day of PEG treatment and the mice regained most of the weight five days after treatment (Fig. 1B). After either PEG, clindamycin, or PEG and clindamycin treatment all mice were challenged with 10^3^ *C. difficile* 630 spores. All treatments rendered mice susceptible to *C. difficile* colonization (Fig. 1C), however PEG-treated mice remained colonized at a high level through thirty days post-infection. In contrast, the clindamycin-treated mice that cleared *C. difficile* within ten days post-infection. Surprisingly, PEG-treated mice were still susceptible to *C. difficile* infection after 10-days of recovery from treatment although *C. difficile* was not detectable in most of the group in the initial five days post-infection (Fig. 1C). From 9 days post-infection onward, the median *C. difficile* stabilized for the 5-day PEG plus 10-day recovery group of mice and remained high through 30 days post-infection (Fig. 1C). Thus, osmotic laxative treatment alone was sufficient to render mice susceptible to prolonged *C. difficile* colonization and PEG-treated mice remained susceptible for up to ten days post-treatment.

@@

```
**5-day laxative treatment differentially disrupts the fecal microbiota compared to clindamycin treatment.** Since laxatives and clindamycin have previously been shown to disrupt the murine microbiota (ref), we hypothesized the different *C. difficile* colonization dynamics between mice treated with laxatives versus clindamycin were due to the two drugs having differential effects on the microbiota. We profiled the stool microbiota over time by sequencing the V4 region of the 16S rRNA gene to compare changes across treatment groups. We found time and treatment group explained half of the observed variation between fecal communities with most of the remaining variation explained by interactions between treatment group and other experimental variables including time, cage effects, and sequencing preparation (PERMANOVA combined R^2^ = `r f_d_top_R2_sum`, *P* `r format.pval(f_d_top_p, eps = 0.001)`, Fig. 2A, Data Set S1, Sheet X). Cage effects refer to the well-documented phenomenon that mice housed in the same cages have similar microbial communities due to copraphagy, we tried to minimize the impact of cage effects on our experiment by breaking up cagemates when assigning mice to treatment groups and primarily housing only two mice per cage. Importantly, although we conducted a total of `r five_day_total_exp` separate experiments, the experiment number and it's interaction with treatment group was not one of the variables that significantly explained the observed variation in fecal communities (Data Set S1, Sheet X). Interestingly, none of the treatment groups recovered to their baseline community structure either 10 or 30 days post-infection suggesting other community features besides recovery to baseline were responsible for the prolonged *C. difficile* colonization in PEG-treated mice (Fig. 2B).

Next, we examined  alpha diversity by looking at Shannon diversity index over time, although both clindamycin and PEG treatments decreased diversity, the diversity index was lower in the groups of mice that received PEG treatment compared to those that received clindamycin through thirty days post-infection (Fig. 2C). We next examined the bacterial genera that shifted after PEG treatment by comparing the baseline samples of mice treated with only PEG to samples from the same mice one day post-treatment. We found `r PEG_genera` OTUs that were altered by PEG, the majority of these OTUs decreased after PEG, but *Enterobacteriaceae* and *Bacteroides* increased and the increase in *Bacteroides* was unique to PEG treatment as Bacteroides actually decreased in clindamycin treated mice (Fig. 2D, Data Set S1, Sheet X). Finally, we examined the bacteria that differ across treatment groups over multiple timepoints. We found `r f_d_group_multi_day` were different over multiple timepoints out of the `r f_d_group_total_genera` that were different between treatment groups (Fig. 2E, Data Set S1, Sheet X). Thus, PEG has a significant impact on the fecal microbiota that was maintained over time and distinct from clindamycin treatment.

Surprisingly, *C. difficile* was not immediately detectable in the stools of the PEG-treated mice that were allowed to recover for 10 days prior to  challenge. We decided to examine the bacteria that changed during the post-infection period when the group median *C. difficile* shifted from undetectable at 1 day post-infection to detectable in the stool samples with the median stabilizing around 8 dayspost-infection (Fig. S1A). Interestingly, we found *Erysipelotrichaceae*, *Enterobacteriaceae*, and *Akkermansia* were changing during the time period when *C. difficile* was becoming detectable in the stools (Fig. S1B), although none of the bacteria were significant after multiple hypothesis correction (Data Set S1, Sheet X). Although we did not identify a clear signal to explain the delayed appearance of *C. difficile* in the 5-day PEG mice that were allowed to recover for 10 days prior to challenge, the delay is striking and could reflect changes in microbial activity or metabolites that were not examined in this study.

@@

#List of variables that significantly explained variation in stool samples
f_d_top_effects_t <- f_d_sign_contrib_t %>% 
  pull(effects)
#Sum of R2 values for variables that significantly explained variation in stool samples
f_d_top_R2_sum_t <- f_d_sign_contrib_t %>% 
  select(R2) %>% sum
#P-value for variables that significantly explained variation in stool samples
f_d_top_p_t <- f_d_sign_contrib_t %>% 
  select(`Pr(>F)`) %>% max # P value will be rounded in text

```
**5-day laxative treatment does not promote more severe CDIs despite altering the mucosal microbiota.** Given the findings from a previous study that demonstrated that PEG treatment disrupts the mucus layer and alters the immune response in mice [@Tropini2018], we decided to examine the impact of PEG treatment on the mucosal microbiota and CDI severity. To evaluate the mucosal microbiota, we sequenced snips of tissue collected from the cecum, proximal colon, and distal colon. Similar to what was observed with the stool samples, alpha diversity was lower in the PEG-treated mice compared to clindamycin treatment (Fig. 3A). Although alpha diversity continued to increase over time based on the communities from PEG-treated mice collected at 20 and 30 days post-infection (Fig. 3A). Group, day, and the interactions with other variables (cage, experiment number, and sample type) explained the majority of the variation observed in mucosal communities (PERMANOVA combined R^2^ = `r f_d_top_R2_sum_t`, *P* `r format.pval(f_d_top_p_t, eps = 0.05)`, Fig. 3B, Data Set S1, Sheet X). *Bacteroides*, *Butyricoccus*, *Clostridiales*, *Clostridium Cluster XIVb*, *Firmicutes*, and *Ruminococcaceae* were consistently different between treatment groups in both the tissue and stool communities (Fig. 3C, Fig. 2E, Data Set S1, Sheet X). Next, we examined CDI severity by evaluating cecum and colon histopathology [@Reeves2011] and found there was no difference in cecum and colon scores between clindamycin and PEG-treated mice that were challenged with *C. difficile* at 4 days post-infection (Fig. 3D), the timepoint typically examined in *C. difficile* 630 challenged mice (ref). We also looked at 6 days post-infection because that was when we started to see a large difference in C*. difficile* colonization levels between PEG- and clindamycin-treated mice (Fig. 1C). Although, there was a slight difference in the colon between PEG and clindamycin-treated mice, there was no difference in the cecum and the overall score is still relatively low given that the max possible summary score is 12 (Fig. 3E). Thus, although PEG treatment had a profound impact on the mucosal microbiota, the impact of *C. difficile* on the cecum and colon was similar between PEG and clindamycin treated mice.

***C. difficile* challenge does not have a synergistic disruptive effect on the microbiota of PEG-treated mice** Because *C. difficile* itself can have an impact on the microbiota [@Jenior2018], we also sequenced the tissue and stools of mock-challenged clindamycin and 5-day PEG treated mice. Examining the stools of the mock-infected PEG- and clindamycin-treated mice revealed similar bacterial disruptions as the *C. difficile* challenged mice (Fig. S2A-C). Similarly, there was no difference between the tissues of mock and *C. difficile* challenged mice (Fig. S2D-F). Thus, most of the microbiota alterations we observed in the PEG-treated mice were a result of the laxative and not an interaction between the laxative and *C. difficile*.

**1-day laxative treatment results in transient *C. difficile* colonization and minor microbiota disruption**

* Draft of results
  + Explain motivation for experiment, set-up of experiment (Fig. 4A). 
  + Transient *C. difficile* colonization in 1-day PEG treated mice with all PEG mice clearing by 7dpi (Fig. 4B)
  + PCoA results (Day has larger R2 than treatment group). PEG mice are closer to baseline communites after 7-day period (Fig. 4C)
  + Shannon results show alpha diversity is only transiently disrupted (Fig. 4D)
  + Highlight bacteria that are disrupted by PEG or clindamycin treatment but recover within 7 days (Fig. 4E)

**Post-CDI laxative treatment disrupts clearance in clindamycin-treated mice regardless of whether an FMT is also administered**

* Draft of results
  + Reiterate motivation for experiment. Saw only a transient microbiota disruption with 1-day PEG treatment alone. Also, see if adding FMT helps with clearance, mention groups have been using PEG or Abx/PEG treatments as a tool to engraft mice with human bacteria communities. Set up of experiment (Fig. 5A).
  + Prolonged *C. difficile* colonization in post-CDI PEG treated mice (Fig. 5B)
  + FMT appears to partially restore alpha diversity (Shannon, but not richness Fig. 5C-D)
  + PCoA (necessary?) Could comment on the clustering of Clindamycin and PEG-treated mice that received FMT (Fig. 5E)
  + Only 2 genera significantly impacted by FMT treatment, likely not as important for clearance (Fig. 5F)
  + Bacteria that consistently differ between groups over time, associated with prolonged colonization (Fig. G)

@@

```
**Five-day post-infection community data can predict mice that will have prolonged *C. difficile* colonization** After identifying bacteria associated with the 5-day, 1-day and post-CDI 1-day PEG treatments, we decided to examine the taxa that were influencing prolonged *C. difficile* colonization. We trained 3 types of machine learning models (random forest, logistic regression, and support vector machine) with input bacterial community data from 5 days post-infection to predict whether the mice were colonized with *C. difficile* 10 days post-infection. We chose 5 days post-infection because that was the earliest timepoint where we would see a treatment effect in the mice that were given 1-day PEG treatment three days after *C. difficile* challenge and then administered an FMT or PBS gavage. The random forest model had the highest performance (AUROC = `r max_auc`, Data Set S1, Sheet X), so we next performed permutation importance to examine the bacteria that were driving performance. We selected the top 20 bacteria contributing to our models performance and examined their relative abundance at 5 days post-infection, the timepoint used to predict C. difficile colonization status on day 10 (Fig. 6A). Next, we focused on the 5 genera that were > 1 % relative abundance in either the cleared or colonized mice and examined the relative abundance dynamics of these bacteria over time. We found _, _, and _ tended to have a higher relative abundance and __, __ had a lower relative abundance in the mice with prolonged colonization compared to the mice that cleared *C. difficile* (Fig. 6B).
Previous work examining the impact of PEG on the murine microbiota found that PEG treatment resulted in the permanent loss of S24-7, also known as *Muribaculum intestinale* [@Tropini2018]. We decided to check our *Porphyromonadaceae* OTUs because *Muribaculum intestinale* is known to be classified as *Porphyromonadaceae* by the Ribosomal Database Project (RDP) database [@Vornhagen2020], *Porphyromonadaceae* was a top feature in the random forest model predicting prolonged *C. difficile* colonization, and had a high relative abundance in the communities of mice that cleared *C. difficile* within 10 days. We identified 4 OTUs that had at least (insert minimum percent identity) to *Muribaculum intestinale* and examined their abundance in mice that either cleared or were still colonized with *C. difficile* at 10 days post-infection (Fig. S3).

@@

* Summary of major findings (Fig. 7A)

@@

* Discuss what these findings might mean for human patients (Fig. 7B)
  + What's known regarding laxatives and susceptibility to CDIs
  + Relevance to human FMTs? Unclear what the best administration route is because there have been no studies designed to evaluate the best administration route for FMTs.

@@

\includegraphics{figure_5.pdf}
A. Experimental timeline for groups receiving PEG treatment post clindamycin gavage. B. Median and individual CFU/g of C. difficile measured over time via culture. Asterisks indicate time points with significant differences between groups using Kruskall-Wallis tests with a Benjamini-Hochberg correction. Limit of detection is 1e2 CFU/g. Background shading indicates PEG treatment period for applicable groups. Opacity of points reflects time point. C. Median and individual Shannon Diversity measured over time. Asterisks, background shading, and opacity of points consistent with (B). D. Median and individual richness measured over time. Asterisks, background shading, and opacity of points consistent with (B).

\newpage
\includegraphics{figure_6.pdf}
**Figure 6. 1-day PEG treatment post C. difficile challenge prolongs colonization regardless of whether an FMT is also administered.** A. PCoA of all groups + FMT plotted over time. Opacity of points reflects time point. B. Relative abundance of genera significantly different over multiple time points post FMT or PBS treatment, plotted over time. Limit of detection is .1%. C. Relative abundances of top 6 significant genera ranked by number of days significant, plotted over time. Limit of detection is .1%.

>>>>

---
title: ''
csl: mbio.csl #Get themes at https://github.com/citation-style-language/styles
fontsize: 11pt
output:
  word_document: default
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
  html_document:
    df_print: paged
geometry: margin=1.0in
bibliography: references.bib
---

@@

**5-day laxative treatment leads to prolonged *C. difficile* colonization in mice.** We compared mice treated with the osmotic laxative PEG 3350 to our standard 10 mg/kg clindamycin treatment, which temporarily renders the mice susceptible to *C. difficile*, with mice typically clearing *C. difficile* within 10 days post-infection [@Tomkovich2019; @Tomkovich2020]. All PEG-treated mice were administered a 15% PEG solution in the drinking water for 5-days, one group was also treated with clindamycin, and one group was allowed to recover for 10 days prior to challenge (Fig. 1A). PEG treatment resulted in weight loss in all 3 groups of PEG-treated mice, with the greatest change in weight observed on the fifth day of PEG treatment and the mice regained most of the weight five days after treatment (Fig. 1B). After either PEG, clindamycin, or PEG and clindamycin treatment all mice were challenged with 10^3^ *C. difficile* 630 spores (Fig. 1A). All treatments rendered mice susceptible to *C. difficile* colonization. However, PEG-treated mice remained colonized with *C. difficile* at high levels through thirty days post-infection (Fig. 1C). In contrast, the clindamycin-treated mice cleared *C. difficile* within ten days post-infection. Surprisingly, PEG-treated mice were still susceptible to *C. difficile* infection after 10-days of recovery from treatment although *C. difficile* was not detectable in most of the group in the initial five days post-infection (Fig. 1C). From 9 days post-infection onward, the median *C. difficile* stabilized for the 10-day recovery group of PEG-treated mice and remained high through 30 days post-infection (Fig. 1C). Thus, osmotic laxative treatment alone was sufficient to render mice susceptible to prolonged *C. difficile* colonization and PEG-treated mice remained susceptible for up to ten days post-treatment.

@@

```
**5-day laxative treatment differentially disrupts the fecal microbiota compared to clindamycin treatment.** Since osmotic laxatives and clindamycin have previously been shown to disrupt the murine microbiota (ref), we hypothesized the different *C. difficile* colonization dynamics between mice treated with the osmotic laxative or clindamycin were due to the two drugs having differential effects on the microbiota. We profiled the stool microbiota over time by sequencing the V4 region of the 16S rRNA gene to compare changes across treatment groups. We found time and treatment group explained half of the observed variation between fecal communities with most of the remaining variation explained by interactions between treatment group and other experimental variables including time, cage effects, and sequencing preparation (PERMANOVA combined R^2^ = `r f_d_top_R2_sum`, *P* `r format.pval(f_d_top_p, eps = 0.001)`, Fig. 2A, Data Set S1, Sheet X). Cage effects refer to the well-documented phenomenon that mice housed in the same cages have similar microbial communities due to copraphagy (ref), we tried to minimize the impact of cage effects on our experiment by breaking up cagemates when assigning mice to treatment groups and primarily housing only two mice per cage. Importantly, although we conducted a total of `r five_day_total_exp` separate experiments, the experiment number and it's interaction with treatment group was not one of the variables that significantly explained the observed variation in fecal communities (Data Set S1, Sheet X). Interestingly, none of the treatment groups recovered to their baseline community structure either 10 or 30 days post-infection suggesting other community features besides recovery to baseline were responsible for the prolonged *C. difficile* colonization in PEG-treated mice (Fig. 2B).

Next, we examined  alpha diversity by looking at Shannon diversity index over time. Although both clindamycin and PEG treatments decreased diversity, Shannon diversity was lower in the groups of mice that received PEG treatment compared to those that received clindamycin through thirty days post-infection (Fig. 2C). We next examined the bacterial genera that shifted after PEG treatment by comparing the baseline samples of mice treated with only PEG to samples from the same mice one day post-treatment. We found `r PEG_genera` genera that were altered by PEG treatment (Data Set S1, Sheet X). The majority of the bacterial relative abundances decreased after PEG, but *Enterobacteriaceae* and *Bacteroides* increased and the increase in *Bacteroides* was unique to PEG treatment as *Bacteroides* actually decreased in clindamycin treated mice (Fig. 2D). Finally, we examined the bacteria that differ across treatment groups over multiple timepoints. We found `r f_d_group_multi_day` genera were different over multiple timepoints out of the `r f_d_group_total_genera` that were different between treatment groups (Fig. 2E, Data Set S1, Sheet X). Thus, PEG has a significant impact on the fecal microbiota that was maintained over time and distinct from clindamycin treatment.

Surprisingly, *C. difficile* was not immediately detectable in the stools of the PEG-treated mice that were allowed to recover for 10 days prior to  challenge. We decided to examine the bacteria that changed during the post-infection period when the group median *C. difficile* shifted from undetectable at 1 day post-infection to detectable in the stool samples with the median stabilizing around 8 days post-infection (Fig. S1A). Interestingly, we found *Erysipelotrichaceae*, *Enterobacteriaceae*, and *Akkermansia* were changing during the time period when *C. difficile* was becoming detectable in the stools (Fig. S1B), although none of the bacteria were significant after multiple hypothesis correction (Data Set S1, Sheet X). Although we did not identify a clear signal to explain the delayed appearance of *C. difficile* in the 5-day PEG mice that were allowed to recover for 10 days prior to challenge, the delay is striking and could reflect changes in microbial activity or metabolites that were not examined in this study.

@@

#List of variables that significantly explained variation in tissue samples
f_d_top_effects_t <- f_d_sign_contrib_t %>% 
  pull(effects)
#Sum of R2 values for variables that significantly explained variation in tissue samples
f_d_top_R2_sum_t <- f_d_sign_contrib_t %>% 
  select(R2) %>% sum
#P-value for variables that significantly explained variation in tissue samples
f_d_top_p_t <- f_d_sign_contrib_t %>% 
  select(`Pr(>F)`) %>% max # P value will be rounded in text

```
**5-day laxative treatment does not promote more severe CDIs despite altering the mucosal microbiota.** Given the findings from a previous study that demonstrated that PEG treatment disrupts the mucus layer and alters the immune response in mice [@Tropini2018], we decided to examine the impact of PEG treatment on the mucosal microbiota and CDI severity. To evaluate the mucosal microbiota, we sequenced snips of tissue collected from the cecum, proximal colon, and distal colon. Similar to what was observed with the stool samples, alpha diversity was lower in the PEG-treated mice compared to clindamycin treatment (Fig. 3A). Although alpha diversity continued to increase over time based on the communities from PEG-treated mice collected at 20 and 30 days post-infection (Fig. 3A, Data Set S1, Sheet X). Group, day of the experiment, and their interactions with other variables (cage, experiment number, and sample type) explained the majority of the variation observed in mucosal communities (PERMANOVA combined R^2^ = `r f_d_top_R2_sum_t`, *P* `r format.pval(f_d_top_p_t, eps = 0.05)`, Fig. 3B, Data Set S1, Sheet X). We saw the greatest difference in the mucosal microbiota between treatment groups at 6 days post-infection with 16 (insert code instead) genera that were significantly different in all three of the tissue types we collected (cecum, proximal colon, and distal colon; Fig. S3A, Data Set S1, Sheet X). However, at 30 days post-infection only *Bacteroides*, *Clostridiales*, *Firmicutes*, and *Ruminococcaceae* were different between treatment groups and just in the cecum tissues (Fig. 3C, Fig. 2E, Data Set S1, Sheet X). Interestingly, one of the genera that was different between treatment groups at 6 days post-infection was *Peptostreptococcaceae* (the genera with sequences that match C. difficile). *Peptostreptococcaceae* was primarily only present in the 5-day PEG treatment group of mice and decreased in the proximal and distal colon tissues over time (Fig. S2B). Thus, PEG treatment had a significant impact on the mucosal microbiota and we detected *C. difficile* sequences in the cecum, proximal colon, and distal colon tissue communities.

Next, we examined CDI severity by evaluating cecum and colon histopathology [@Reeves2011] and found there was no difference in cecum and colon scores between clindamycin and PEG-treated mice that were challenged with *C. difficile* at 4 days post-infection (Fig. 3D), the timepoint typically examined in *C. difficile* 630 challenged mice (ref). We also looked at 6 days post-infection because that was when we started to see a large difference in C*. difficile* colonization levels between PEG- and clindamycin-treated mice (Fig. 1C). Although, there was a slight difference in the colon between PEG and clindamycin-treated mice, there was no difference in the cecum and the overall score is still relatively low given that the max possible summary score is 12 (Fig. 3E). Thus, although PEG treatment had a profound impact on the mucosal microbiota, the impact of *C. difficile* 630 infection on the cecum and colon was similar between PEG and clindamycin treated mice.

***C. difficile* challenge does not have a synergistic disruptive effect on the microbiota of PEG-treated mice** Because *C. difficile* itself can have an impact on the microbiota [@Jenior2018], we also sequenced the tissue and stools of mock-challenged clindamycin and 5-day PEG treated mice. Examining the stools of the mock-challenged mice revealed similar bacterial disruptions as the *C. difficile* challenged mice (Fig. S2A-C). Similarly, there was no difference between the tissues of mock and *C. difficile* challenged mice (Fig. S2D-F). Thus, most of the microbiota alterations we observed in the PEG-treated mice were a result of the laxative and not an interaction between the laxative and *C. difficile*.
``` {r 1-day PEG stool microbiota}

#Genera that were different between treatment days
#Genera that were changed by treatment
pt_d1 <- read_tsv(here("data/process/1_Day_PEG_genus_stats_PTtoD1.tsv")) %>% 
  filter(p.value.adj < 0.05) 
#Genera that were still different at 7 days post-infection  
pt_d7 <- read_tsv(here("data/process/1_Day_PEG_genus_stats_PTtoD7.tsv")) %>% 
  filter(p.value.adj < 0.05)
pt_d7_genera <- pt_d7 %>% pull(genus)

recovered_genera <- pt_d1 %>% 
  filter(!genus %in% pt_d7_genera) %>% #Remove genera that were still different at day 7
  nrow() #Tally number of rows

```
**1-day laxative treatment results in transient *C. difficile* colonization and minor microbiota disruption** Next, we decided to examine how a shorter osmotic laxative perturbation would impact the microbiome and susceptibility to *C. difficile*. We administered mice either a 1-day PEG treatment, a 1-day PEG treatment with a 1-day recovery period, or a clindamycin treatment before challenging them with *C. difficile 630* (Fig. 3A). In contrast to the 5-day PEG treated mice, the 1-day PEG treated mice were only transiently colonized and cleared *C. difficile* by 7 days post-infection (Fig. 3B). The stool communities of PEG-treated mice were also only transiently disrupted, with Shannon diversity close to baseline by 7 days post-infection (Fig. 3C-D). We found `r recovered_genera` bacteria were impacted by treatment but recovered close to baseline levels within 7 days post-infection including *Enterobacteriaceae*, *Clostridiales*, *Porpyromonadaceae*, and *Ruminococcaceae* (Fig. 3E, Data Set S1, Sheet X). These findings suggest the duration of treatment matters when considering the impact of the laxatives, with shorter treatments resulting in a transient loss of *C. difficile* colonization resistance.

**Post-CDI laxative treatment disrupts clearance in clindamycin-treated mice regardless of whether an FMT is also administered** Since 1-day PEG treatment resulted in a more mild microbiota perturbation, we decided to use the 1-day treatment to examine the hypothesis that PEG helps to flush *C. difficile* spores from the intestine. We were also interested in exploring whether PEG might help with engraftment in the context of FMTs. To examine the impact of PEG treatment on *C. difficile* clearance and FMT treatment, we treated 4 groups of mice with clindamycin and then challenged all mice with *C. difficile* before administering the following treatments: no additional treatment, 1-day PEG immediately after challenge, and 1-day PEG treatment 3 days after challenge followed by either administration of an FMT or PBS solution via gavage (Fig. 5A). In contrast to our hypothesis, all groups of mice that received PEG actually exhibited prolonged *C. difficile* colonization (Fig. 5B).

* Draft of results

  + Prolonged *C. difficile* colonization in post-CDI PEG treated mice (Fig. 5B)
  + FMT appears to partially restore alpha diversity (Shannon, but not richness Fig. 5C-D)
  + PCoA (necessary?) Could comment on the clustering of Clindamycin and PEG-treated mice that received FMT (Fig. 6A)
  + Only 2 genera significantly impacted by FMT treatment, likely not as important for clearance (Fig. 6B)
  + Bacteria that consistently differ between groups over time, associated with prolonged colonization (Fig. 6C)

@@

```
**Five-day post-infection community data can predict mice that will have prolonged *C. difficile* colonization** After identifying bacteria associated with the 5-day, 1-day and post-CDI 1-day PEG treatments, we decided to examine the taxa that were influencing prolonged *C. difficile* colonization. We trained 3 types of machine learning models (random forest, logistic regression, and support vector machine) with input bacterial community data from 5 days post-infection to predict whether the mice were colonized with *C. difficile* 10 days post-infection. We chose 5 days post-infection because that was the earliest timepoint where we would see a treatment effect in the mice that were given 1-day PEG treatment three days after *C. difficile* challenge and then administered an FMT or PBS gavage. The random forest model had the highest performance (AUROC = `r max_auc`, Data Set S1, Sheet X), so we next performed permutation importance to examine the bacteria that were driving performance. We selected the top 10 bacteria contributing to our models performance (Fig. 7A) and examined their relative abundance at 5 days post-infection, the timepoint used to predict C. difficile colonization status on day 10 (Fig. 7B). Next, we focused on the 5 genera that were > 1 % relative abundance in either the cleared or colonized mice and examined the relative abundance dynamics of these bacteria over time. We found _, _, and _ tended to have a higher relative abundance and __, __ had a lower relative abundance in the mice with prolonged colonization compared to the mice that cleared *C. difficile* (Fig. 7C).
Previous work examining the impact of PEG on the murine microbiota found that PEG treatment resulted in the permanent loss of S24-7, also known as *Muribaculum intestinale* [@Tropini2018]. We decided to check our *Porphyromonadaceae* OTUs because *Muribaculum intestinale* is known to be classified as *Porphyromonadaceae* by the Ribosomal Database Project (RDP) database [@Vornhagen2020], *Porphyromonadaceae* was a top feature in the random forest model predicting prolonged *C. difficile* colonization, and had a high relative abundance in the communities of mice that cleared *C. difficile* within 10 days. We identified 4 OTUs that had at least (insert minimum percent identity) to *Muribaculum intestinale* and examined their abundance in mice that either cleared or were still colonized with *C. difficile* at 10 days post-infection (Fig. S4).

@@

* Summary of major findings (Fig. 8A)

@@

* Discuss what these findings might mean for human patients (Fig. 8B)
  + What's known regarding laxatives and susceptibility to CDIs
  + Relevance to human FMTs? Unclear what the best administration route is because there have been no studies designed to evaluate the best administration route for FMTs.

@@

\includegraphics{figure_5.pdf}
**Figure 5. 1-day PEG treatment post *C. difficile* challenge prolongs colonization regardless of whether an FMT is also administered.** A. Experimental timeline for groups receiving PEG treatment post clindamycin gavage. B. Median and individual CFU/g of C. difficile measured over time via culture. Asterisks indicate time points with significant differences between groups using Kruskall-Wallis tests with a Benjamini-Hochberg correction. Limit of detection is 1e2 CFU/g. Background shading indicates PEG treatment period for applicable groups. Opacity of points reflects time point. C. Median and individual Shannon Diversity measured over time. Asterisks, background shading, and opacity of points consistent with (B). D. Median and individual richness measured over time. Asterisks, background shading, and opacity of points consistent with (B).

\newpage
\includegraphics{figure_6.pdf}
**Figure 6. For 1-day PEG treatment post *C. difficile* challenge mice that also receive an FMT only some bacterial genera were restored.** A. PCoA of all groups + FMT plotted over time. Opacity of points reflects time point. B. Relative abundance of genera significantly different over multiple time points post FMT or PBS treatment, plotted over time. Limit of detection is .1%. C. Relative abundances of top 6 significant genera ranked by number of days significant, plotted over time. Limit of detection is .1%.
